---
title: "Card Sorting Report"
author: "Hampus Lennaárd - Data Scientist @ Cohr Group"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(jsonlite)
library(purrr)
library(papaja)
library(knitr)
options(digits = 3)
```

```{r load-data, echo=FALSE, warning=FALSE, message=FALSE}
card_data_unclean <- read.csv("E:/thesis_2024/Data_bucket/thesis_raw_card_data.csv") %>%
  rename(user_id = test_taker_id) # renaming due to script being written with another id column used 

# Adding a names_df to keep track of ids 
names_df <- card_data_unclean %>%
  distinct(user_id, .keep_all = TRUE) %>%
  select(user_id, firstname, lastname)

total_sample_size <- length(unique(card_data_unclean$user_id))
```

### Overview of Card Sorting Test 
The total sample contained `r total_sample_size` test-takers.

```{r process_questions, echo=FALSE, warning=FALSE, message=FALSE}
questions_df <- card_data_unclean %>%
  select(question_id, question_correct_answer, question_configuration) %>%
  distinct(question_id, .keep_all = TRUE)

# Function to parse JSON safely, seperating the concatenated strings 
parse_json_safe <- function(json_str) {
  tryCatch(
    fromJSON(json_str),
    error = function(e) NA  # Return NA if parsing fails
  )
}

# Apply the function to each row in the question_configuration column
questions_df <- questions_df %>%
  mutate(parsed_config = map(question_configuration, parse_json_safe)) %>%
  select(-question_configuration)

find_rule <- function(questions_df) {
  
  # Helper function to determine rule of current card
    find_common_value <- function(correct_card, objective) {
      for(col in names(correct_card)) {
        if(correct_card[[col]] == objective[[col]]) {
          last_common_col <- col
        }
      }
      return(last_common_col) # returning last col to avoid instances were > 1 common cols are found, see note at top of report
    }
  
  
  for(i in 1:nrow(questions_df)) {
    # Fetch the correct answer
    correct_answer <- questions_df$question_correct_answer[i]

    # Fetch the json list for manipulation
    json_list <- questions_df$parsed_config[[i]]
    
    # Fetch card-options
    card_options <- json_list$options

    # Determine which card was correct & fetch the data for that card 
    correct_card <- card_options[correct_answer,1:3]
    
    # Fetch data for sorting card
    objective <- as.data.frame(json_list$objective) %>%
      select(-c("position", "backgroundColor")) # removing columns that could mess with "shapes" and "shapeColor" when searching for rule
  
    # Determine rule
    rule <- find_common_value(correct_card, objective)
    
    questions_df$rule[i] <- rule # creating the new "rule" column in df and assigning current rule of question to question row

  }
    return(questions_df)
}

questions_final_df <- find_rule(questions_df)

questions_final_df <- questions_final_df %>%
  mutate(
    rule_number = case_when(
      rule == "shapeColor" ~ 1,
      rule == "shapes" ~ 2,
      rule == "shape" ~ 3,
      TRUE ~ NA_real_  # fail-safe
    )
  )

# Finally join questions_final_df with the card_data_unclean 
card_data_expanded <- inner_join(card_data_unclean, questions_final_df, by = "question_id") %>%
  select(-c("question_correct_answer.y", "question_configuration")) %>% # cleaning up some side effects
  rename(question_correct_answer = question_correct_answer.x) 
```

```{r clean_card_data, echo=FALSE, warning=FALSE, message=FALSE}

# Helper function to check if a test-taker answer is repeated consecutively for at least 20 times (indicating guessing only)
has_consecutive_repeats <- function(data, n = 10) {
  rle_vec <- rle(data$test_taker_answer)
  if (any(rle_vec$lengths >= n)) {
    return(TRUE)
  }
  return(FALSE)
}

# Below code is commented out because data-processing has already been done in this case 

# Assuming card_data_expanded is already defined
# Create a dataframe with test_taker_ids that have consecutive repeats
#consecutive_repeats_df <- card_data_expanded %>%
#  group_by(test_taker_id) %>%
#  filter(has_consecutive_repeats(cur_data(), 10)) %>%
#  ungroup()

# Create a cleaned dataframe excluding those test_taker_ids with consecutive repeats
#card_data_clean <- card_data_expanded %>%
#  group_by(test_taker_id) %>%
#  filter(n() > 20) %>%
#  filter(!is.na(test_taker_answer)) %>%
#  ungroup()

# Temp to avoid restructuring
card_data_clean <- card_data_expanded %>%
 
total_sample <- length(unique(card_data_clean$test_taker_id))
```

In order to clean the data-set I decided to filter out people who had answered less than 20 questions in total, counting these people as faulty trails. Moreover a simple function was written to filter away people who had answered the same card consecutively for 10 questions in a row. Indicating that at some point the test-taker stopped conducting the test properly and started just re-pressing the same card over and over again, regardless of the active rule. 

This resulted in a total N of `r total_sample`

```{r function-categorize_responses, echo=FALSE, warning=FALSE, message=FALSE}
categorise_responses <- function(dataframe) {
  
  unique_ids <- unique(dataframe$user_id)

  results <- tibble(
    user_id = unique_ids,
    corr_ans = integer(length(unique_ids)),
    p_error = integer(length(unique_ids)),
    non_p_error = integer(length(unique_ids)),
    trails_first_category = integer(length(unique_ids)),
    t_error = integer(length(unique_ids)),
    questions_administrated = integer(length(unique_ids)),
    conceptual_level_responses = integer(length(unique_ids))
  )
  
  for (id in unique_ids) {
    id_data <- dataframe %>%
      filter(user_id == id) %>%
      arrange(question_order) # ensure rows are arranged from 1:100 / last
    
    shift_questions <- c(6, 11, 16, 21, 26, 31, 36, 41, 46, 51, 56, 61, 66, 71, 76, 81, 86, 91, 96)
    
    # Initialize relevant variables
    p_error <- 0
    non_p_error <- 0
    corr_ans <- 0
    t_error <- 0
    first_category_complete <- FALSE
    consecutive_logic <- FALSE
    consecutive_correct <- 0
    
    rules_applied_vec <- c()
    
    for (question in 1:nrow(id_data)) {
      current_rule <- id_data$rule_number[question]
      correct_answer <- id_data$question_correct_answer[question]
      test_taker_answer <- id_data$test_taker_answer[question]
      
      if (is.na(test_taker_answer)) { # NA present breaks the response-pattern completely 
            break # Exit the inner loop and move to the next id
      }  
      
      # Figuring out which rule was applied:
      
      # Step 1 - Fetch card details 
      card_details <- id_data$parsed_config[[question]]$options
    
      # Step 2 - Fetch the card test-taker chose 
      chosen_card_details <- card_details[test_taker_answer, ]
      
      # Step 3 - Fetch the details of the card to be sorted 
      objective_card_details <- as.data.frame(id_data$parsed_config[[question]]$objective) %>%  # its a list so convert to df
        select(-c(position, backgroundColor)) # remove these columns as they can mess with evaluation of commonalities

      # Step 4 - Compare both cards to find commonalities
      common_feature <- which(chosen_card_details == objective_card_details)
      
     # Step 5 - Ensure to get only the first common feature (one card has >1 common features)
      if (length(common_feature) > 0) {
        common_feature <- common_feature[1]
      } else {
        common_feature <- NA
      }

      # Step 6 - Decide which rule was applied 
      rule_used <- if (!is.na(common_feature)) names(chosen_card_details)[common_feature] else NA # setting rule used to 0 for ambigious sorts - these are exploration errors with no clear hypothesis from test-taker 

      # Step 7 - Convert to numeric so we can compare with rule_number
      if (!is.na(rule_used)) {
        rule_applied <- switch(rule_used,
                               "shapeColor" = 1,
                               "shapes" = 2,
                               "shape" = 3,
                               0) # default case, ambiguous sort - exploration with no clear hypothesis from the test-taker
      } else {
        rule_applied <- 0
      }
      
      rules_applied_vec[question] <- rule_applied
      
      if (question == 1) {
        if (correct_answer == test_taker_answer) {
          corr_ans <- corr_ans + 1
          consecutive_logic <- TRUE
        } else {
          non_p_error <- non_p_error + 1
          t_error <- t_error + 1
          consecutive_logic <- FALSE
        }
        if (first_category_complete == FALSE && correct_answer == test_taker_answer) {
          results$trails_first_category[results$user_id == id] <- question
          first_category_complete <- TRUE
        }
      } else {
        previous_rule <- id_data$rule_number[question - 1]
        previous_applied_rule <- rules_applied_vec[question - 1]

        if (correct_answer == test_taker_answer) {
          corr_ans <- corr_ans + 1
          if (consecutive_logic) {
            consecutive_correct <- consecutive_correct + 1
          }
          consecutive_logic <- TRUE
        } else {
          t_error <- t_error + 1
          consecutive_logic <- FALSE
          
          if (question %in% shift_questions) {
            non_p_error <- non_p_error + 1
          } else {
            if (previous_rule == current_rule && rule_applied == previous_applied_rule) {
              p_error <- p_error + 1
            } else {
              non_p_error <- non_p_error + 1
            }
          }
        }
        if (first_category_complete == FALSE && correct_answer == test_taker_answer) {
          results$trails_first_category[results$user_id == id] <- question
          first_category_complete <- TRUE
        }
      }
    }
    
    results$corr_ans[results$user_id == id] <- corr_ans
    results$p_error[results$user_id == id] <- p_error
    results$non_p_error[results$user_id == id] <- non_p_error
    results$t_error[results$user_id == id] <- t_error
    results$questions_administrated[results$user_id == id] <- corr_ans + t_error
    results$conceptual_level_responses[results$user_id == id] <- consecutive_correct
  }
  
  return(results)
}


```
```{r call-categorise_responses, echo=FALSE, warning=FALSE, message=FALSE}
final_card_data <- categorise_responses(card_data_clean)

# Using time_used from the original "results" dataframe, time_used in this dataframe is not capped at 150 seconds so time is incorrect

#Joining final data with names, writing to excel
final_card <- inner_join(names_df, final_card_data, by = "user_id")
writexl::write_xlsx(final_card, "E:/thesis_2024/Data_bucket/final_card_data.xlsx")

# Calculating proportions
card_proportions <- final_card %>%
  summarise(
    total_p_error = sum(p_error, na.rm = TRUE),
    total_non_p_error = sum(non_p_error, na.rm = TRUE),
    total_trails_first = sum(trails_first_category, na.rm = TRUE),
    total_clr = sum(conceptual_level_responses, na.rm = TRUE),
    total_questions = sum(questions_administrated, na.rm = TRUE),
    mean_prop_p_error = total_p_error / total_questions,
    mean_prop_non_p_error = total_non_p_error / total_questions,
    mean_prop_trails_first = total_trails_first / total_questions,
    mean_prop_clr = total_clr / total_questions
  )

writexl::write_xlsx(card_proportions, "E:/thesis_2024/Data_bucket/card_proportions.xlsx")
```

### Summary Statistics of final data 
**NOTE** 

"clr" : Conceptual Level Responses

"p_error": perseverative errors

"non_p_error": non-perseverative errors

```{r summarise_test_results, echo=FALSE, warning=FALSE, message=FALSE}
summary_stats_correct <- final_card_data %>%
  summarise(
    Mean_Correct = mean(corr_ans),
    SD_Correct = sd(corr_ans),
    Mean_trails_first = mean(trails_first_category),
    SD_trails_first = sd(trails_first_category),
    Mean_clr = mean(conceptual_level_responses),
    SD_clr = sd(conceptual_level_responses)
  )

kable(summary_stats_correct, caption = "Summary: Correct Answers & Trails to first category")

summary_stats_errors <- final_card_data %>%
  summarise(
    Mean_p_error = mean(p_error),
    SD_p_error = sd(p_error),
    Mean_non_p_error = mean(non_p_error),
    SD_non_p_error = sd(non_p_error)
  )

kable(summary_stats_errors, caption = "Summary: Type of Errors")



proportions <- final_card_data %>%
  summarise(
    mean_prop_p_error = mean(p_error/questions_administrated),
    mean_prop_non_p_error = mean(non_p_error/questions_administrated),
    mean_prop_trails_first = mean(trails_first_category/questions_administrated),
    mean_prop_t_error = mean(t_error/questions_administrated),
    mean_prop_clr = mean(conceptual_level_responses/questions_administrated),
    mean_prop_corr = mean(corr_ans/questions_administrated)
  )
```

#### Total Correct Answers
In WCST the total correct categories achieved measures cognitive flexibility where higher numbers are desirable. 
```{r plot-correct_answers, echo=FALSE}
bins_use <- round(sqrt(length(final_card_data$corr_ans)))
 
ggplot(final_card_data, aes(x = corr_ans)) +
  geom_histogram(bins = bins_use, alpha = 0.6, colour = "blue", fill = "blue") +
  annotate("text", x = 70, y = 30, label = "Mean = 38", color = "black") +
  annotate("text", x = 70, y = 25, label = "SD = 18", color = "black") +
  theme_apa() +
  labs(x = "Total Correct Answers", y = "Total Count", title = "Total Correct Answers")

```

#### Trails to Complete First Category
In WCST the trails to complete first category reflects initial cognitive flexibility and the test-takers ability to generate and test new solutions to faced problems. A lower number is desirable, indicating greater initial flexibility and problem-solving ability. 
```{r plot-trails_first_category, echo=FALSE}
bins_use <- round(sqrt(length(final_card_data$trails_first_category)))
 
ggplot(final_card_data, aes(x = trails_first_category)) +
  geom_histogram(bins = bins_use, alpha = 0.6, colour = "blue", fill = "blue") +
  annotate("text", x = 15, y = 75, label = "Mean = 3.96", color = "black") +
  annotate("text", x = 15, y = 65, label = "SD = 3.01", color = "black") +
  theme_apa() +
  labs(x = "Trails to 1st Correct Sort", y = "Total Count", title = "Trails to First Correct Sort")

```

#### Perseverative Errors
In WCST the perseverative errors are errors where the test-taker continues to use a previously correct rule despite feedback indicating that it is no longer corrrect. The measure thus reflect difficulties in shifting cognitive strategies. A lower number is therefore desirable, indicating better ability to shift cognitive strategies. 

```{r plot-p_errors, echo=FALSE}
bins_use <- round(sqrt(length(final_card_data$p_error)))

ggplot(final_card_data, aes(x = p_error)) +
  geom_histogram(bins = bins_use, alpha = 0.6, colour = "blue", fill = "blue") +
  annotate("text", x = 55, y = 30, label = "Mean = 32.5", color = "black") +
  annotate("text", x = 55, y = 25, label = "SD = 13.1", color = "black") +
  theme_apa() +
  labs(x = "Perseverative Errors", y = "Total Count", title = "Perseverative Errors")

```

#### Non-Perseverative Errors
In WCST, the Non-Perseverative Errors are errors where the test-taker makes an error that is not based on the previous rule, it therefore reflects guessing / seeking behaviour and can be seen as a proxy for ones ability to generate and test new solutions to problems. 
```{r plot-non_p_errors, echo=FALSE}
bins_use <- round(sqrt(length(final_card_data$non_p_error)))
 
ggplot(final_card_data, aes(x = non_p_error)) +
  geom_histogram(bins = bins_use, alpha = 0.6, colour = "blue", fill = "blue") +
  annotate("text", x = 25, y = 45, label = "Mean = 14.6", color = "black") +
  annotate("text", x = 25, y = 40, label = "SD = 4.13", color = "black") +
  theme_apa() +
  labs(x = "Non-Perseverative Errors", y = "Total Count", title = "Non-Perseverative Errors")

```

#### Conceptual Level Responses
In WCST, the conceptual level responses measures the number of consecutive correct responses after a test-taker has sorted a card correctly, it reflects, on a conceptual-level the test-takers ability to identify, learn and apply the rule consistently. It is therefore a measure of several aspects of cognitive flexibility, predominantly sustained attention as the test-taker has to be able to keep the rule in mind and apply it consecutively and error monitoring / correction as the test-taker has to recognize when the rule is incorrect, figure out the new rule and then consistently apply it once more. A higher number is thus desirable. 
```{r plot-conceptual_level_responses, echo=FALSE}
bins_use <- round(sqrt(length(final_card_data$conceptual_level_responses)))

ggplot(final_card_data, aes(x = conceptual_level_responses)) +
  geom_histogram(bins = bins_use, alpha = 0.6, colour = "blue", fill = "blue") +
  annotate("text", x = 55, y = 25, label = "Mean = 23.9", color = "black") +
  annotate("text", x = 55, y = 20, label = "SD = 14.5", color = "black") +
  theme_apa() +
  labs(x = "CLR", y = "Total Count", title = "Conceptual Level Responses")

```

**Final Note** 
All bins were calculated according to the "square-root choice method": 
number_of_bins = sqrt(length(final_card_data$x))