---
title: "Color Word Outcome Variables - Report"
author: "Hampus Lennaárd - Data Scientist @ Cohr Group"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(jsonlite)
library(purrr)
library(papaja)
library(knitr)
options(digits = 3)
```

```{r load-data, echo=FALSE, warning=FALSE, message=FALSE}
color_word_data_unclean <- read.csv("raw_color_data.csv")

total_sample_size <- length(unique(color_word_data_unclean$test_taker_id))
```

### Overview of Color Word Test 
The total sample contained `r total_sample_size` test-takers.

```{r process_questions, echo=FALSE, warning=FALSE, message=FALSE}
# Subset a DF to only include once instance of each question 
questions_df <- color_word_data_unclean %>%
  select(question_id, question_configuration) %>%
  distinct(question_id, .keep_all = TRUE)

# Function to parse JSON safely, seperating the concatenated strings 
parse_json_safe <- function(json_str) {
  tryCatch(
    fromJSON(json_str),
    error = function(e) NA  # Return NA if parsing fails
  )
}

# Apply the function to each row in the question_configuration column
questions_df <- questions_df %>%
  mutate(parsed_config = map(question_configuration, parse_json_safe))

# For-loop to replace "position" into an index and "objective" into a workable format 
pos_to_index <- function(config) {
  pos <- config$position
  
  pos_index <- switch(pos,
                      "top left" = 1,
                      "top middle" = 2,
                      "top right" = 3,
                      "middle left" = 4,
                      "middle middle" = 5,
                      "middle right" = 6,
                      "bottom left" = 7,
                      "bottom middle" = 8,
                      "bottom right" = 9,
                      NA)
  
  config$objective_position <- pos_index
  config$objective <- ifelse(config$objective == "color of the text", "color", "text")
  return(config)
}

# Apply the function to each parsed_config
questions_df <- questions_df %>%
  mutate(parsed_config = map(parsed_config, pos_to_index))

# Function to determine if a question is congruent or incongruent
is_congruent <- function(questions_df) {

  for (i in 1:nrow(questions_df)) {
    config <- questions_df$parsed_config[[i]]
    
    obj_index <- config$objective_position
    
    text <- config$options$text[obj_index]
    color <- config$options$color[obj_index]
      
    if (text == color) {
      question_type <- "congruent"
      } else {
        question_type <- "incongruent"
      }
    
    questions_df$question_type[i] <- question_type
  }
  
  questions_df <- questions_df %>%
    select(question_id, question_type)
  
  return(questions_df)
}

# Apply the function to update the dataframe
questions_final_df <- is_congruent(questions_df)

congruent_questions_df <- questions_final_df %>%
  filter(question_type == "congruent")

total_congruent_questions <- nrow(congruent_questions_df)

incongruent_questions_df <- questions_final_df %>%
  filter(question_type == "incongruent")

total_incongruent_questions <- nrow(incongruent_questions_df)
```

The Color Word Test includes `r total_congruent_questions` congruent questions and `r total_incongruent_questions` incongruent questions. 

```{r process_color_data, echo=FALSE, warning=FALSE, message=FALSE}
color_word_data_unclean <- color_word_data_unclean %>%
  group_by(test_taker_id) %>%
  mutate(
    id = group_indices()
  ) %>%
  ungroup() %>%
  select(
    id, question_id, question_correct_answer, test_taker_answer, time_used
  ) %>%
  mutate(
    time_used = time_used * 0.001 # converting time_used from miliseconds to seconds 
  )

# summary(color_word_data_unclean)

color_word_data_clean <- color_word_data_unclean %>%
  group_by(id) %>%
  filter(n() > 20) %>%
  filter(time_used < 46) %>% # filtering time_used below 46 seconds, anything above is due to low MS at test-taker level, is 46 to high even? 
  ungroup()

# summary(color_word_data_clean)

total_n <- length(unique(color_word_data_clean$id))

color_word_data_clean <- inner_join(color_word_data_clean, questions_final_df, by = "question_id")

color_word_incongruent <- color_word_data_clean %>%
  filter(question_type == "incongruent") %>%
  mutate(
    correct = ifelse(question_correct_answer == test_taker_answer, 1, 0),
    incongruent_error = ifelse(question_type == "incongruent" & correct == 0, 1, 0),
    incongruent_rt = ifelse(question_type == "incongruent" & correct == 1, time_used, 0)
  )

color_word_congruent <- color_word_data_clean %>%
  filter(question_type == "congruent") %>%
  mutate(
    correct = ifelse(question_correct_answer == test_taker_answer, 1, 0),
    congruent_error = ifelse(question_type == "congruent" & correct == 0, 1, 0),
    congruent_rt = ifelse(question_type == "congruent" & correct == 1, time_used, 0)
  )

color_word_data_full <- color_word_data_clean %>%
  mutate(
    correct = ifelse(question_correct_answer == test_taker_answer, 1, 0),
    congruent_error = ifelse(question_type == "congruent" & correct == 0, 1, 0),
    congruent_rt = ifelse(question_type == "congruent" & correct == 1, time_used, 0),
    incongruent_error = ifelse(question_type == "incongruent" & correct == 0, 1, 0),
    incongruent_rt = ifelse(question_type == "incongruent" & correct == 1, time_used, 0)
  )

color_word_final <- color_word_data_full %>%
  group_by(id) %>%
  summarise(
    questions_administrated = n(),
    congruent_error_rate = sum(congruent_error, na.rm = TRUE) / questions_administrated,
    congruent_rt = mean(congruent_rt, na.rm = TRUE),
    incongruent_error_rate = sum(incongruent_error, na.rm = TRUE) / questions_administrated,
    incongruent_rt = mean(incongruent_rt, na.rm = TRUE),
    total_correct = sum(correct, na.rm = TRUE)
  )

# Summary statistics for congruent data
congruent_summary <- color_word_final %>%
  summarise(
    Mean_ER = mean(congruent_error_rate, na.rm = TRUE),
    Sd_ER = sd(congruent_error_rate, na.rm = TRUE),
    Mean_RT = mean(congruent_rt, na.rm = TRUE),
    Sd_RT = sd(congruent_rt, na.rm = TRUE)
  )

# Summary statistics for incongruent data
incongruent_summary <- color_word_final %>%
  summarise(
    Mean_ER = mean(incongruent_error_rate, na.rm = TRUE),
    Sd_ER = sd(incongruent_error_rate, na.rm = TRUE),
    Mean_RT = mean(incongruent_rt, na.rm = TRUE),
    Sd_RT = sd(incongruent_rt, na.rm = TRUE)
  )

proportions <- color_word_final %>%
  summarise(
    mean_prop_inccongruent_error_rate = mean(incongruent_error_rate/questions_administrated),
    mean_prop_incongruent_rt = mean(incongruent_rt/questions_administrated),
    mean_prop_congruent_error_rate = mean(congruent_error_rate/questions_administrated),
    mean_prop_congruent_rt = mean(congruent_rt/questions_administrated),
  )

write.csv(proportions, "color_word_props.csv")
```

### Summary Statistic Congruent Questions
```{r print-congruent_summary, echo=FALSE}
kable(congruent_summary, caption = "Summary Statistics for Congruent Data")
```

#### Congruent plots
```{r plot-congruent_errors, echo=FALSE, results='asis'}

con_er_bins <- round(sqrt(length(color_word_final$congruent_error_rate)))
 
ggplot(color_word_final, aes(x = congruent_error_rate)) +
  geom_histogram(bins = con_er_bins, alpha = 0.6, colour = "blue", fill = "blue") +
  annotate("text", x = 0.3, y = 100, label = "Mean = 0.015", color = "black") +
  annotate("text", x = 0.3, y = 85, label = "SD = 0.040", color = "black") +
  theme_apa() +
  labs(x = "Error Rate(ER)", y = "Total Count", title = "Congruent Error Rate Distribution")

con_rt_bins <- round(sqrt(length(color_word_final$congruent_rt)))

ggplot(color_word_final, aes(x = congruent_rt)) +
  geom_histogram(bins = con_rt_bins, alpha = 0.6, colour = "blue", fill = "blue") +
  annotate("text", x = 4.5, y = 45, label = "Mean = 3.04", color = "black") +
  annotate("text", x = 4.5, y = 40, label = "SD = 0.768", color = "black") +
  theme_apa() +
  labs(x = "Reaction Time (RT)", y = "Total Count", title = "Congruent RT Distribution")

```

### Final Summary statistics for the incongruent condition
```{r print-incongruent_summary, echo=FALSE}
kable(incongruent_summary, caption = "Summary Statistics for Incongruent Data")
```

```{r plot-incongruent_errors, echo=FALSE, results='asis'}

incon_er_bins <- round(sqrt(length(color_word_final$incongruent_error_rate)))


ggplot(color_word_final, aes(x = incongruent_error_rate)) +
  geom_histogram(bins = incon_er_bins, alpha = 0.6, colour = "blue", fill = "blue") +
  annotate("text", x = 0.4, y = 60, label = "Mean = 0.046", color = "black") +
  annotate("text", x = 0.4, y = 50, label = "SD = 0.071", color = "black") +
  theme_apa() +
  labs(x = "Error Rate (ER)", y = "Total Count", tite = "Incongruent Error Rate Distribution")


incon_rt_bins <- round(sqrt(length(color_word_final$incongruent_rt)))


ggplot(color_word_final, aes(x = incongruent_rt)) +
  geom_histogram(bins = incon_rt_bins, alpha = 0.6, colour = "blue", fill = "blue") +
  annotate("text", x = 7.5, y = 50, label = "Mean = 4.01", color = "black") +
  annotate("text", x = 7.5, y = 40, label = "SD = 1.02", color = "black") +
  theme_apa() +
  labs(x = "Reaction Time (RT)", y = "Total Count", title = "Incongruent RT Distribution")

```

#### Extra: Paired T-Test between The two Conditions (RT)
```{r t_test_rt, echo=FALSE, warning=FALSE, message=FALSE}
# Since I plotted RT for both conditions and they were normally distributed I will not perform a shapiro test 
res <- t.test(color_word_final$congruent_rt, color_word_final$incongruent_rt, paired = TRUE)
```

A paired samples t.test was performed on the reaction time of the congruent vs incongruent conditions, the result provide strong evidence for the existance of a "stroop-interference effect" in Cohr Group's Color Word Test, t(`r res$parameter`) = `r res$statistic`, p = < .001. CI [`r res$conf.int[1]`, `r res$conf.int[2]`]. Overall, the t.test resulted in a sample estimate mean difference = `r res$estimate`